---
title: "Data Manipulation"
author: "Steph Locke"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document: 
    code_folding: show
    number_sections: yes
    toc: yes
    toc_float: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Data Manipulation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

[Home](index.html) | [Prev](GettingStarted.html) | [Next]()

# Working with vectors {.tabset}
## Intro
Vectors are the building blocks of R. They are 1D arrays that contain data of all the same type - so coercion of data types happens.

- R is a 1-based array system so a vector starts at position 1
- Items in a vector can have names but can also be referred to by position

## Create
You can create vectors using the `c()` function or via a sequence (`:`). Values in an array can be given names - this can be useful when subsetting values/
```{r}
handCrafted<-c(1,2,3,4)
seqCrafted<-1:4
named<-c(a=1,b=2,c=3,d=4)
named
```


## Filter
Alternatively called subsetting we can filter a vector by using positive notations of position, negative notations of position, the name of a value, or providing a boolean value for each item in the vector.
```{r}
handCrafted[1]
seqCrafted[-1]
named["b"]
# Here I use shorthands for TRUE and FALSE for brevity - not generally recommended
handCrafted[c(T,T,T,F)]
```

## Update
You can update one or more values in a vector by assigning the new values into the desired subset.
```{r}
handCrafted
handCrafted[2]<-99
handCrafted

named
named["a"]<-99
named

# Delete by subsetting without value
seqCrafted
seqCrafted<-seqCrafted[-4]
seqCrafted
```

## Manipulate
You can manipulate a vector using functions. You can overwrite any object created - this will change mode, class, etc as R is **loosely typed** and doesn't require such things to be specified and fixed up front.
```{r}
mode(seqCrafted)
seqCrafted<-as.character(seqCrafted)
mode(seqCrafted)
```

## Order
Ordering of records works by providing the position numbers of values in a vector and then using those to produce a vector with the original components in new locations
```{r}
preOrder<-sample(letters, 6)
preOrder

# Get order the values should appear in to be alphabetised
order(preOrder)

# Use it to sort a vector
ordered<-preOrder[order(preOrder)]
ordered

# Alternatively, use the sort() function for brevity
sorted<-sort(preOrder)
sorted
```

## Metadata
You can extract various pieces of information about a vector
```{r}
names(handCrafted)
names(named)

dim(named)
dimnames(named)

length(named)
class(named)
mode(named)
attributes(named)
```

# Working with lists

# Working with tables

# Get data from SQL Server
