---
title: "Data Manipulation"
author: "Steph Locke"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document: 
    code_folding: show
    number_sections: yes
    toc: yes
    toc_float: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Data Manipulation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

[Home](index.html) | [Prev](GettingStarted.html) | [Next]()

# Working with vectors {.tabset}
## Intro
Vectors are the building blocks of R. They are 1D arrays that contain data of all the same type - so coercion of data types happens.

- R is a 1-based array system so a vector starts at position 1
- Items in a vector can have names but can also be referred to by position

## Create
```{r}
handCrafted<-c(1,2,3,4)
seqCrafted<-1:4
named<-c(a=1,b=2,c=3,d=4)

```

## Update
```{r}
handCrafted
handCrafted[2]<-99
handCrafted

named
named["a"]<-99
named
```

## Manipulate
```{r}
mode(seqCrafted)
seqCrafted<-as.character(seqCrafted)
mode(seqCrafted)
```

## Order
```{r}
preOrder<-sample(letters, 6)
preOrder

# Get order values should appear in
order(preOrder)
# Use it
ordered<-preOrder[order(preOrder)]


# Can also sort
sorted<-sort(preOrder)
sorted
```

## Filter


# Working with lists

# Working with tables

# Get data from SQL Server
