---
title: "Calling SQL Server from R and vice versa"
author: "Steph Locke"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document: 
    code_folding: show
    number_sections: yes
    toc: yes
    toc_float: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Calling SQL Server from R and vice versa}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

[Home](index.html) | [Prev](DataViz.html) | [Next]()

# Calling SQL Server from R {.tabset}

## Building a connection
- Need a driver name!
- Need credentials (you can do trusted AD connections too)
- Need to store the connection for use
```{r}
library(RODBC)
#host<-"rmsftdp.northeurope.cloudapp.azure.com"
host<-"104.45.80.173"
db<-"AWDW16"
uname<-"Rdemo"
pwd<-"tugait"
connstring<-sprintf("DRIVER={ODBC Driver 11 for SQL Server};Server=%1$s;Database=%2$s;Uid=%3$s;Pwd=%4$s;",host,db,uname,pwd)

conn<-odbcDriverConnect(connstring)
# Test!
if(conn!=-1L) sqlTables(conn,schema = "dbo")
```

## RODBC functions
RODBC comes with some built-in functions for working with your database
```{r eval=FALSE}
dt<- sqlFetch(conn, "DatabaseLog")
sqlCopy( conn, "select * from DatabaseLog", "DatabaseLog2")
sqlSave( conn, dt[1:100,], "DatabaseLog2")
sqlClear(conn, "DatabaseLog2")
sqlDrop( conn, "DatabaseLog2")
```

## SQL
Use normal SQL statements using `sqlQuery()` - must send one statement at a time to get the results back from each.

```{r}
dt<- sqlQuery(conn,"
         SELECT 
          DatabaseLogID
          ,DatabaseUser
          ,Event
          ,[Schema]
          ,Object
          FROM DatabaseLog
         ")
knitr::kable(head(dt))
```


```{r}
str( sqlQuery(conn,"
        SELECT COUNT(*) as Vol FROM DatabaseLog;
         SELECT DatabaseLogID FROM DatabaseLog;
         ")
)
```

# Calling R from SQL Server {.tabset}
## Executing R
Executing R from SQL Server requires the execution of a stored procedure called `sp_execute_external_script` that takes three main parameters:

- `@language` This gets set to R but isn't isn't it an interesting parameter with lots of possibilities for the future?
- `@script` This is the R code to be executed
- `@input_data_1` This is a SQL statement that will produce the starting dataset in R

Unfortunately, the return schema is not inherited by SQL Server so you (ideally) should provide an output specification  via `WITH RESULT SETS`
```{r echo=FALSE}
# convenience function
sqlServer<-function(sql) sqlQuery(conn, sql)
```


```{r}
sqlServer(
"EXECUTE   sp_execute_external_script
                @language = N'R'
              , @script = N'OutputDataSet <- InputDataSet'
              , @input_data_1 = N'SELECT 1 as Col'
 WITH RESULT SETS ((col int not null))           "
)
```

## Limitations
- Output schema not inherited
- Single table passthrough
- Just one output

## Installation notes
[BOL](https://msdn.microsoft.com/en-us/library/mt604845.aspx)
