---
title: "Calling SQL Server from R"
author: "Steph Locke"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document: 
    code_folding: show
    number_sections: yes
    toc: yes
    toc_float: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Calling SQL Server from R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

[Home](index.html) | [Prev](DataViz.html) | [Next]()

# Calling SQL Server from R {.tabset}

## Building a connection
- Need a driver name!
- Need credentials (you can do trusted AD connections too)
- Need to store the connection for use
```{r}
library(RODBC)
host<-"Insufficientdata.northeurope.cloudapp.azure.com"
db<-"AdventureWorks2016CTP3"
uname<-"Rdemo"
pwd<-"SQLbits16!"
connstring<-sprintf("DRIVER={ODBC Driver 11 for SQL Server};Server=%1$s;Database=%2$s;Uid=%3$s;Pwd=%4$s;",host,db,uname,pwd)

conn<-odbcDriverConnect(connstring)
# Test!
if(conn!=-1L) sqlTables(conn,schema = "dbo")
```

## RODBC functions
RODBC comes with some built-in functions for working with your database
```{r eval=FALSE}
dt<- sqlFetch(conn, "DatabaseLog")
sqlCopy( conn, "select * from DatabaseLog", "DatabaseLog2")
sqlSave( conn, dt[1:100,], "DatabaseLog2")
sqlClear(conn, "DatabaseLog2")
sqlDrop( conn, "DatabaseLog2")
```

## SQL
Use normal SQL statements using `sqlQuery()` - must send one statement at a time to get the results back from each.

```{r}
dt<- sqlQuery(conn,"
         SELECT 
          DatabaseLogID
          ,DatabaseUser
          ,Event
          ,[Schema]
          ,Object
          FROM DatabaseLog
         ")
knitr::kable(head(dt))
```


```{r}
str( sqlQuery(conn,"
        SELECT COUNT(*) as Vol FROM DatabaseLog;
         SELECT DatabaseLogID FROM DatabaseLog;
         ")
)
```
