---
title: "Calling R from SQL Server"
author: "Steph Locke"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document: 
    code_folding: show
    toc: yes
    toc_float: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Calling R from SQL Server}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

[Home](index.html) | [Prev](SQLServerFromR.html) | [Next]()

## Executing R
Executing R from SQL Server requires the execution of a stored procedure called `sp_execute_external_script` that takes three main parameters:

- `@language` This gets set to R but isn't isn't it an interesting parameter with lots of possibilities for the future?
- `@script` This is the R code to be executed
- `@input_data_1` This is a SQL statement that will produce the starting dataset in R

Unfortunately, the return schema is not inherited by SQL Server so you (ideally) should provide an output specification  via `WITH RESULT SETS`
```{r echo=FALSE}
library(RODBC)
#host<-"rmsftdp.northeurope.cloudapp.azure.com"
host<-"104.45.80.173"
db<-"AWDW16"
uname<-"Rdemo"
pwd<-"tugait"
connstring<-sprintf("DRIVER={ODBC Driver 11 for SQL Server};Server=%1$s;Database=%2$s;Uid=%3$s;Pwd=%4$s;",host,db,uname,pwd)

conn<-odbcDriverConnect(connstring)
# convenience function
sqlServer<-function(sql) sqlQuery(conn, sql)
```


```{r}
sqlServer(
"EXECUTE   sp_execute_external_script
                @language = N'R'
              , @script = N'OutputDataSet <- InputDataSet'
              , @input_data_1 = N'SELECT 1 as Col'
 WITH RESULT SETS ((col int not null))           "
)
```

## Limitations
- Output schema not inherited
- Single table passthrough
- Just one output

## Installation notes
[BOL](https://msdn.microsoft.com/en-us/library/mt604845.aspx)
