---
title: "R and Microsoft"
author: "Steph Locke"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
vignette: >
  %\VignetteIndexEntry{R and Microsoft}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# R
> R is an integrated suite of software facilities for data manipulation, calculation and graphical display

- Open source
- In-memory & single-core (by default)
- Multi-platform
- Extensible environment
- Delivered by the [R Foundation](), supported by the [R Consortium](), grown by R developers 
- [r-project.org](http://www.r-project.org/)

## There's a package for that
- R is extended with packages of functionality
- More than 8,000 active packages according to [MetaCRAN](http://www.r-pkg.org/)
- [Comprehensive R Archive Network (CRAN)](https://cran.r-project.org/)
- [Github](https://github.com) via [devtools](https://cran.r-project.org/package=devtools)

## How it hangs together
```{r,echo=FALSE,warning=FALSE}
library(DiagrammeR)
hangs<-"graph TB
subgraph High Level
D[Data] --> R[Read into memory]
R --> DM[Data manipulation]
DM --> V[Visualisation]
DM --> S[Statistical Analysis]
DM --> O[Output]
S --> V
V --> Do[Document]
Do --> O
V --> I[Interactive report]
I --> O
S --> W[Webservice]
W --> I
W --> O
end
subgraph Package View
Dp[RDBMS, flat file, nosql, xml] --> Rp[data.table, RODBC, httr, rvest, readxl, readr,...]
Rp --> DMp[data.table, dplyr, zoo, lubridate, stringr, ...]
DMp --> Vp[ggplot2, DiagrammeR, vegalite, ...]
DMp --> Sp[caret, survival, nnet, ...]
DMp --> Op[base, readr, RODBC, ...]
Sp --> Vp
Vp --> Dop[knitr, rmarkdown, ...]
Dop --> Op
Vp --> Ip[shiny, ...]
Ip --> Op
Sp --> Wp[AzureML, jug, ...]
Wp --> Ip
Wp --> Op
end"
mermaid(hangs)
```

# Microsoft Data Platform

## How it hangs togethers
```{r,echo=FALSE,warning=FALSE}
library(DiagrammeR)
hangs<-"graph TB
subgraph High Level
D[Data] --> R[ETL]
R --> DM[Database]
DM --> S[Cube]
DM --> I[Itneractive reports]
DM --> Do[Static reports]
DM --> W[Webservices]
DM --> O[Output]
S --> I
S --> Do
S --> O
I --> O
Do --> O
W --> I
W --> O
end
subgraph Software View
Dm[Data] --> Rm[SSIS, Data Factory]
Rm --> DMm[SQL Server, Azure SQL]
DMm --> Sm[SSAS]
DMm --> Im[PowerBI]
DMm --> Dom[SSRS]
DMm --> Wm[AzureML]
DMm --> Om[Output]
Sm --> Im
Sm --> Dom
Sm --> Om
Im --> Om
Dom --> Om
Wm --> Im
Wm --> Om
Wm --> Om
end"
mermaid(hangs)
```
